Bioverse: Potentially Observable Exoplanet Biosignature Patterns Under the UV Threshold Hypothesis for the Origin of Life

### 2025-02-05 src/tex/ms.tex
add statistical constraints to intro

\subsection{Semi-analytical approach}\label{sec:met-semianalytical}
%First, we apply a Bayes Factor Design Analysis~\citep{Schoenbrodt2018}
\rev{In order to constrain the evidence yield of potential future surveys aimed at evaluating the} UV Threshold Hypothesis ($H_1$) above, \rev{we first derive a simplified analytical approach.
This approach ignores the effects of exoplanet demographics and sample selection to focus on the qualitative effects of sample size and the unknown probablity of abiogenesis on the evidence yield of future surveys.
To do so, we need to frame hypotheses $H_\mathrm{null}$ and $H_1$ as probabilistic models that determine the probability distributions of outcomes in terms of samples of exoplanets for which biosignatures are detected or not.}
%We first assessed the expected probabilities of obtaining true negative or true positive evidence for the UV Threshold Hypothesis ($H_1$) above, as well as the probability for misleading or inconclusive evidence, under idealized conditions.
%This serves as a first-order estimate of the information content of a survey, before we take into account the effects of exoplanet demographics, sample selection, and survey strategy.

@@

\rev{Next, we determine the probability distribution of sample outcomes, or likelihood of each hypothesis.}
Let $Y=\sum_i^n L_i$ be the random variable counting the number of positive life detections in a sample of size $n$. Its probability mass function under the null hypothesis $H_\mathrm{null}$ is that of a binomial distribution:
\begin{equation}
    \label{eq:semian:likelihoodHnull}
P(Y=k|H_\mathrm{null}) = \binom{n}{k}f_\mathrm{life}^k(1-f_\mathrm{life})^{n-k}.
\end{equation}

@@

\rev{Then, we turn to a metric which measures the information gain of sampling: the Bayes factor \cite{Jeffreys1939}.
The posterior evidence for the null hypothesis $H_{\mathrm{null}}$ is given by the Bayes' theorem:
\begin{equation}
\label{eq:bayesHnull}
    P(H_{\mathrm{null}}|Y) = \frac{P(Y|H_{\mathrm{null}})P(H_{\mathrm{null}})}{P(Y)}
\end{equation}
and the posterior evidence for the NUV hypothesis:
\begin{equation}
    \label{eq:bayesHone}
        P(H_1|Y) = \frac{P(Y|H_1)P(H_1)}{P(Y)}~.
\end{equation}
Computing the distributions given by equations \ref{eq:bayesHnull} and \ref{eq:bayesHone} requires the distributions of likelihoods, given in equations \ref{eq:semian:likelihoodHnull} and \ref{eq:semian:likelihoodH1}, as well as the prior distributions for each hypothesis ($P(H_{\mathrm{null}})$ and $P(H_{1})$) and for the sample ($P(H_{1})$).
}
\rev{
Apart from the caveat that the likelihood for $H_{1}$ cannot be expressed analytically due to being conditioned by n_{\lambda} (equation \ref{eq:semian:likelihoodH1}), computing posterior probabilities is challenging due to the unknown priors.
}

@@

\rev{Our Bayes factor analysis quantifies how much evidence future data may provide in favor of a hypothesis against another one (the null hypothesis).
Typical frequentist (p-value based) tests usually evaluate only the likelihood of the null hypothesis, thus requiring that the experiment be carefully crafted such that rejection of the null hypothesis can be interpreted in favor of a particular studied hypothesis (for instance that samples from coin tosses score a very low likelihood under the hypothesis of a fair coin may indicate that the coin is biased).
In contrast, Bayes factor analyses such as ours introduce additional constraint to the testing by involving the likelihood of the tested alternate hypothesis (equation \ref{eq:bayes_factor}).
Thus, under Bayes-factor analyses, evidence for the tested hypothesis is not only measured by how unlikely the null hypothesis is given some observation but rather by how much more likely than the null hypothesis is the alternate hypothesis.
The most exhaustive form of hypothesis selection test would be performed by using the entire Bayes theorem (equation \ref{eq:BayesTheorem}), computing the \emph{posterior} probability of the tested hypothesis.
Doing so enables to quantify the information gain from performing some observation (by comparing the posterior with the prior), and to perform inference with consideration for prior information.
\begin{equation}
    \label{eq:BayesTheorem}
    P(H_i|X) = \frac{P(X|H_i)P(H_i)}{P(X)}
\end{equation}
However, implementing the Bayes theorem can prove difficult in practice, owing to the (usually) subjective nature of prior distributions.
Two prior distributions are required for the computation of the posterior distribution: the prior distribution for the considered model and hypothesis, and the prior distribution for the observation.
While the former requires that at least some amount of consensus is identifiable in the community of scientists that discuss this hypothesis, the latter also requires that either empirical distributions for the prior distribution of the observation be available or that the exhaustive list of mechanistic pathways that could potentially lead to this observation be known and tractable.
Taking the ratio of the posterior evidence for two alternate hypotheses (i.e. of equation \ref{eq:BayesTheorem}) allows this problematic term, the prior probability for the observation, to cancel out (equation \ref{eq:postratio}).
\begin{equation}
    \label{eq:postratio}
    \frac{P(H_1|X)}{P(H_0|X)} = \frac{P(X|H_0)P(H_0)}{P(X|H_0)P(H_0)}
\end{equation}
This trades off the opportunity to estimate the posterior probability of a specific hypothesis with the possiblity to narrow down the problem to two specific competing hypotheses.
Could there be factors other than the past UV flux that would determine the occurrence rate of biosignatures in terrestrial exoplanets?
Surely, there is a vast array of possible mechanisms that could be considered which could shape the prior probability of an observed sample of biosignatures $X$.
Truncating the ratio of prior probabilities from equation \ref{eq:postratio} results in the Bayes factor (equation \ref{eq:bayes_factor}), which we used in our semi-analytical analysis (section \ref{sec:results-semianalytical}).
While the Bayes ratio only paints part of the picture of what a full inference pipeline would look line in a future survey aimed at testing origin of life hypotheses from the observation of biosignatures, it captures an essential part of the yield of a particular strategy: by how much would these observations tip the scale of model selection relative to prior expectations.}

\rev{Our Bayes factor analysis quantifies how much evidence future data may provide in favor of a hypothesis against another one (the null hypothesis).
Typical frequentist (p-value based) tests usually evaluate only the likelihood of the null hypothesis, thus requiring that the experiment be carefully crafted such that rejection of the null hypothesis can be interpreted in favor of a particular studied hypothesis (for instance that samples from coin tosses score a very low likelihood under the hypothesis of a fair coin may indicate that the coin is biased).
In contrast, Bayes factor analyses such as ours introduce additional constraint to the testing by involving the likelihood of the tested alternate hypothesis (equation \ref{eq:bayes_factor}).
Thus, under Bayes-factor analyses, evidence for the tested hypothesis is not only measured by how unlikely the null hypothesis is given some observation but rather by how much more likely than the null hypothesis is the alternate hypothesis.
The most exhaustive form of hypothesis selection test would be performed by using the entire Bayes theorem (equation \ref{eq:BayesTheorem}), computing the \emph{posterior} probability of the tested hypothesis.
Doing so enables to quantify the information gain from performing some observation (by comparing the posterior with the prior), and to perform inference with consideration for prior information.
\begin{equation}
    \label{eq:BayesTheorem}
    P(H_i|X) = \frac{P(X|H_i)P(H_i)}{P(X)}
\end{equation}
However, implementing the Bayes theorem can prove difficult in practice, owing to the (usually) subjective nature of prior distributions.
Two prior distributions are required for the computation of the posterior distribution: the prior distribution for the considered model and hypothesis, and the prior distribution for the observation.
While the former requires that at least some amount of consensus is identifiable in the community of scientists that discuss this hypothesis, the latter also requires that either empirical distributions for the prior distribution of the observation be available or that the exhaustive list of mechanistic pathways that could potentially lead to this observation be known and tractable.
Taking the ratio of the posterior evidence for two alternate hypotheses (i.e. of equation \ref{eq:BayesTheorem}) allows this problematic term, the prior probability for the observation, to cancel out (equation \ref{eq:postratio}).
\begin{equation}
    \label{eq:postratio}
    \frac{P(H_1|X)}{P(H_0|X)} = \frac{P(X|H_0)P(H_0)}{P(X|H_0)P(H_0)}
\end{equation}
This trades off the opportunity to estimate the posterior probability of a specific hypothesis with the possiblity to narrow down the problem to two specific competing hypotheses.
Could there be factors other than the past UV flux that would determine the occurrence rate of biosignatures in terrestrial exoplanets?
Surely, there is a vast array of possible mechanisms that could be considered which could shape the prior probability of an observed sample of biosignatures $X$.
Truncating the ratio of prior probabilities from equation \ref{eq:postratio} results in the Bayes factor (equation \ref{eq:bayes_factor}), which we used in our semi-analytical analysis (section \ref{sec:results-semianalytical}).
While the Bayes ratio only paints part of the picture of what a full inference pipeline would look line in a future survey aimed at testing origin of life hypotheses from the observation of biosignatures, it captures an essential part of the yield of a particular strategy: by how much would these observations tip the scale of model selection relative to prior expectations.}

>>>>

\subsection{Semi-analytical approach}\label{sec:met-semianalytical}
%First, we apply a Bayes Factor Design Analysis~\citep{Schoenbrodt2018}
We first assessed the expected probabilities of obtaining true negative or true positive evidence for the UV Threshold Hypothesis ($H_1$) above, as well as the probability for misleading or inconclusive evidence, under idealized conditions.
This serves as a first-order estimate of the information content of a survey, before we take into account the effects of exoplanet demographics, sample selection, and survey strategy.

@@

Let $Y=\sum_i^n L_i$ be the random variable counting the number of positive life detections in a sample of size $n$. Its probability mass function under the null hypothesis $H_\mathrm{null}$ is that of a binomial distribution:
\begin{equation}
P(Y=k|H_\mathrm{null}) = \binom{n}{k}f_\mathrm{life}^k(1-f_\mathrm{life})^{n-k}.
\end{equation}

@@

@@
